Language Attitudes of Catalan-Spanish Bilinguals Towards Intervocalic Alveolar Fricatives in Catalan

### 2024-04-29 manuscript/main.tex

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
  a4paper,
  11pt,
  twocolumn]{article}
\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math} % this also loads fontspec
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else
  % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{48,48,48}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{1.00,0.81,0.69}{#1}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.50,0.62,0.50}{\textbf{#1}}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.80,0.80,0.80}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.86,0.64,0.64}{#1}}
\newcommand{\BuiltInTok}[1]{\textcolor[rgb]{0.80,0.80,0.80}{#1}}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.86,0.64,0.64}{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.50,0.62,0.50}{#1}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.50,0.62,0.50}{\textbf{#1}}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.86,0.64,0.64}{\textbf{#1}}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.94,0.87,0.69}{#1}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.87,0.87,0.75}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.86,0.86,0.80}{#1}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.50,0.62,0.50}{#1}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.76,0.75,0.62}{#1}}
\newcommand{\ExtensionTok}[1]{\textcolor[rgb]{0.80,0.80,0.80}{#1}}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.75,0.75,0.82}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.94,0.94,0.56}{#1}}
\newcommand{\ImportTok}[1]{\textcolor[rgb]{0.80,0.80,0.80}{#1}}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.50,0.62,0.50}{\textbf{#1}}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.94,0.87,0.69}{#1}}
\newcommand{\NormalTok}[1]{\textcolor[rgb]{0.80,0.80,0.80}{#1}}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.94,0.94,0.82}{#1}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.94,0.94,0.56}{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{1.00,0.81,0.69}{\textbf{#1}}}
\newcommand{\RegionMarkerTok}[1]{\textcolor[rgb]{0.80,0.80,0.80}{#1}}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.86,0.64,0.64}{#1}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.80,0.58,0.58}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.80,0.58,0.58}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.80,0.80,0.80}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.80,0.58,0.58}{#1}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.50,0.62,0.50}{\textbf{#1}}}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
%------------------------------------------------------------------------------%
% PAPER TEMPLATE FOR ICPHS 2023 Prague                                         %
%                                                                              %
% Original template downloaded from:                                           %
% http://www.icphs2023.org/call-for-papers/                                    %
%                                                                              %
% Reformatted to work with Rmarkdown and R by:                                 %
% Joseph V. Casillas | Rutgers Univesity |11/11/2022                           %
%                                                                              %
% Available for download at:                                                   %
% https://github.com/jvcasillas/icphs2023_rmd_template                         %
%------------------------------------------------------------------------------%

@@

\title{Paper template for {ICPhS} 2023 Prague}
\author{Please write XXX instead of the name(s) of the author(s)}
\organization{Please write XXX instead of the affiliation(s)}
\email{please write XXX instead of the email address(es)}

@@

\begin{abstract}
This is the layout specification and template definition for the papers of ICPhS XX (the 20th International Congress of Phonetic Sciences, which will be held in Prague, Czech Republic, August 7-11, 2023). 
This template is a revised version of the one used for Melbourne ICPhS XIX in 2019, originally generated from the template for Speech Prosody 2006 in Dresden.

The abstract may consist of more than one paragraph but must be kept within a 150-word limit. 
This abstract will be printed in the abstract booklet to be given out at the conference.
\end{abstract}

\keywords{There is space for up to five self-selected keywords (maximally two 
lines).}

@@

Catalan and Spanish are two languages that have coexisted for centuries,
mainly in eastern Spain, along the Mediterranean cost \textbf{(Hualde,
1992)}. It is safe to assume that there is a bidirectional influence
between these languages: in that region, Catalan has influenced Spanish
and Spanish has influenced Catalan in all linguistic levels
\textbf{(Davidson, 2014, 2015)}. A crucial area of cross-linguistic
influence might be the production of sounds, especially in those sounds
who have different distribution in Catalan and Spanish. For example,
previous studies have found that, in Barcelona, the production of
Catalan vowels is different based on the L1 of the speaker, being it
less target-like in L1 Spanish speakers compared to L1 Catalan speakers
\textbf{(CITAR SEBASTIAN GALLES, etc)}.

As for how Catalan people perceive these differences in production and,
thus, their attitudes towards a less Catalan-like production of a given
phoneme little is known. The aim of this study is to examine the covert
attitudes towards the production of intervocalic /s/, as either {[}s{]}
(disfavored by prescriptivist norms) or {[}z{]} (preferred by
prescriptivist norms) by L1 Catalan, L1 Spanish, and Catalan-Spanish
bilinguals from Catalonia using a matched guise test \textbf{(Lambert,
1967)}. The alveolar fricative was chosen because it exists in both
Catalan and Spanish inventories, but their contextual distribution is
different (Llisterri, 2023).

@@

Interestingly, Medieval Spanish, similarly to Modern Catalan,
distinguished between /s/ and /z/ in intervocalic position within a
word, although these sounds merged through a devoicing process
\cite{hualde_colina_2015}. Nevertheless, in those modern varieties where
the /s/ in coda position is produced, such as in Peninsular Spanish,
this voiceless phoneme might assimilate to the following sound,
especially if it is a voiced consonant (\emph{mismo}, {[}mízmo{]}),
rather than a vowel (\emph{las alas}, {[}lasálas{]}) \textbf{(Hualde \&
Colina, 2015; Hualde et al., 2020)}. Hence, Spanish inventory does not
have the phoneme /z/, but {[}z{]} may occur as an allophone of /s/, due
to a voice assimilation process. It is important to mention, though,
that it has been claimed that this voicing is generally not complete,
but partial \textbf{(Benet et al., 2012; Hualde \& Colina, 2015; Hualde
et al., 2020)}.

There is some evidence of the influence of Spanish on the production of
the Catalan alveolar fricative, in contexts where, according to
prescriptivist rules \textbf{(Institut d'Estudis Catalans, 2022)}, the
{[}z{]} should be produced, but {[}s{]} is produced instead. This
phenomenon has mainly been attested in those areas where there has been
an extended contact between Catalan and Spanish, such as the so-called
``franja de Aragón'' (borderland between Catalonia and Aragon), some
areas of the Valencian Community, and Barcelona and other urban areas
\textbf{(Benet et al., 2012; Davidson, 2020; Recasens Vives, 1996,
p.~268)}. The opposite pattern has also been observed - that is, the
voicing of the alveolar fricative in the Spanish spoken in Barcelona and
Valencia. Interestingly, although in both Barcelona and Valencia Catalan
and Spanish coexist and influence each other, it seems that Catalan has
a greater influence on Spanish in Barcelona, but Spanish has a greater
influence on Catalan in Valencia \textbf{(Davidson, 2020)}.

In summary, both Catalan and Spanish have a voiceless alveolar fricative
in their inventories. Catalan also has a voiced variant, which is
produced when this sound is followed by a voiced consonant or a vowel.
Spanish might produced this sound only when it is followed by a voiced
consonant, although /s/ might be ``partially voiced'' (Hualde \& Colina,
2015). Thus, these phonemes, although they have the same allophones,
their contextual distribution is different (Llisterri, 2023). Moreover,
some studies have shown that there is a bidirectional influence between
both languages in the production of the voiced vs the voiceless alveolar
fricative, and that such influence might be different based on the
context. These divergent patterns might be explained by the different
sociopolitic and linguistic realities of each city, and might have an
effect on speakers' attitudes towards the devoicing of the alveolar
fricative /z/ in Catalan.

@@

The social meaning and the perception of intervocalic {[}z{]} in Spanish
has been studied in a variety of contexts, including both monolingual
\textbf{(Chappell, 2016, 2017, 2021; Chappell et al., 2023; Chappell \&
García, 2017; García, 2015, 2019; Silverstein, 2003)} and bilingual
contexts \textbf{(Davidson, 2014, 2015, 2019; Sinner, 2002)}, despite
having been categorized as a low-salient variant \textbf{(Chappell,
2017, 2021; Chappell et al., 2023; García, 2015)}. With respect to the
former, it has been found that for listeners from San José, Costa Rica,
the covert attitudes associated with intervocalic {[}z{]} vary
significantly by speaker gender. Consequently, male speakers who use
{[}z{]} tend to be positively evaluated, and generally perceived as
nice, local, confident, and masculine. Conversely, female speakers who
use {[}z{]} are only afforded negative associations of lower education
and social class \textbf{(Chappell, 2016; Chappell et al., 2023;
Silverstein, 2003)}.

Similarly, \textbf{García (2019)} shows that, in Ecuador, intervocalic
/s/ is evaluated differently in male and female speakers. Accordingly,
women who produce intervocalic {[}z{]} are viewed as significantly lower
status, less pleasant, and older. Contrary, men who produce intervocalic
{[}z{]} do not receive these negative evaluations. Additionally, this
variable has been found to serve as a marker of regional identity, with
voicing being associated with neighboring Highland Ecuadorian dialects
(Cuenca and Quito) that exhibit systematic voicing. \textbf{García
(2020)} hypothesizes that young men are the leaders of this change,
essentially due to the fact that they are in greater contact with these
highland varieties. Interestingly, the mean voicing rates of females are
roughly two age groups behind those of males. Consequently, it is
believed that it is not necessarily the case that women in Ecuador are
avoiding voicing like San José women due to its negative evaluations,
but rather that they are further behind in the adoption of the incoming
variant.

With respect to the social meaning and the perception of intervocalic
{[}z{]} in bilingual contexts, this phenomenon has been mainly studied
in Catalan-Spanish contact situations, given its robust usage by
Catalan-Spanish bilinguals and, in particular, by those with greatest
exposure to and usage of Catalan \textbf{(Davidson, 2014, 2015)}.
According to \textbf{Sinner (2002)}, intervocalic {[}z{]} can be
considered a possible linguistic marker of Catalonian Spanish, despite
its relative lack of social awareness \textbf{(Davidson, 2019; Sinner,
2002)}. In their study, Sinner (2002) interviewed 12 speakers of
Barcelonan Spanish and monolingual Spanish-speakers from Madrid aged 27
to 41 regarding their awareness of linguistic features of Catalanized
Spanish. Results showed that only select Barcelona speakers perceived
intervocalic {[}z{]} as a linguistic marker of Catalonian Spanish,
whereas monolingual Spanish-speakers from Madrid did not. Additionally,
Davidson (2019) found that, in Barcelona, the indexical field of
intervocalic {[}z{]} features exclusively positive associations with
bilingualism and local solidarity, which warrants its gradual adoption
in the speech community, unlike other phonetic markers of Catalonian
Spanish that are the topic of overt social ridicule, such as lateral
velarization. These results are in line with previous findings on
intervocalic {[}z{]} in Catalonian Spanish. For instance,
\textbf{Silverstein (2003)} found that intervocalic {[}z{]} indexes a
local Barcelona identity, and the variant then acquires positive
associations with bilingualism and cultural pride. Overall, these
findings suggest that intervocalic {[}z{]} likely functions as a
sociolinguistic indicator of Catalonian Spanish within the Catalonian
speech community, indexing Catalan-Spanish bilingual membership while
lacking any considerable degree of generalized conscious awareness
\textbf{(Davidson, 2019)}.

@@

\textbf{INTRO}. Accordingly, the research questions and hypotheses are
as follows: 1. What covert attitudes are linked to the perception
Catalan intervocalic /s/ as either {[}s{]} or {[}z{]}? 2. What is the
relationship between individual social factors (mother tongue, province
of origin) and covert attitudes towards Catalan intervocalic /s/?

@@

Data were collected using an online survey hosted on Qualtrics that
consisted of two parts. The first part, a demographic questionnaire,
collected information regarding participant age, gender, place of
origin, place of residence, mother tongue, family income, and level of
studies. The second portion of the survey consisted of a perception task
where participants were presented with a series of stimuli containing
the target phoneme contrast (intervocalic voiced and voiceless alveolar
fricative) followed by 11 evaluative statements related to social
characteristics such as level of studies, friendliness, social class,
and ethnic origin. These evaluative statements were then followed by a
100-point draggable sliding scale with which participants indicated
their level of agreement or disagreement with the content of the
statement, where 0 was ``totally disagree'' and 100 was ``totally
agree''.

@@

Data were analyzed in R (R Core Team, 2023). First, we performed a
descriptive analysis that consisted of creating a series of plots and
tables to summarize and describe the features of the dataset and to
identify potential trends. After that, we performed inferential
statistical analysis. For each of the 11 evaluative statements, we
created a series of hierarchically nested linear regression models using
the lme4 package (Bates et al., 2015) with attitudinal sliding scale
responses as the dependent variable, and phoneme, mother tongue, and
province as potential predictors. To determine the best fit model for
each statement, we used the anova() function (Fox \& Weisberg, 2019).

@@

Participant responses to the statements \emph{Parla bé català} `S/he
speaks proper Catalan' and \emph{És d'origen català} `S/he is of Catalan
origin' show similar results, as illustrated in Figures \ref{fig:pbc}
and \{fig:oc\}. Those who have both Catalan and Spanish as their L1 rate
the speaker more positively regardless of phoneme, compared to those
whose L1 is either Catalan or Spanish. In addition, although it is not
statistically significant, participants give higher ratings to the
speaker when they are listening to those statements with {[}z{]}.

\begin{figure}[!ht]
\begin{center}
\includegraphics[height=15cm]{./includes/figures/pbc.png}
\caption{Ratings of "Parla bé català" as a function of mother tongue}\label{fig:pbc}
\end{center}
\end{figure}

@@

Participant responses to the statements \emph{És simpàtic} `S/he is
nice' and \emph{És de fiar} `S/he is trustworthy' showed similar
results, as illustrated in Figures \ref{fig:simpatic} and \{fig:fiar\}.
Those who are from and live in the province of Barcelona give higher
ratings to the speaker, compared to those who are from and live in
Girona.

\begin{figure}[!ht]
\begin{center}
\includegraphics[height=15cm]{./includes/figures/simpatic.png}
\caption{Ratings of "És simpàtic" as a function of mother tongue}\label{fig:simpatic}
\end{center}
\end{figure}

@@

\section{Discussion and Conclusions}

The aim of this exploration study was to investigate whether phoneme
production ({[}s{]} vs {[}z{]}), mother tongue (Catalan vs Spanish vs
both), and province of origin (Girona vs Barcelona) influenced the
attitudes of the participants towards a native speaker of Central
Catalan. The results suggest that these predictors have a different
effect on the ratings of the speaker depending on the social
characteristic under evaluation.

@@

\subsubsection{Headings}

Section headings are centered in boldface with capitalized letters.
Sub-headings start at the left margin in the column with the first
letter capitalized and the rest of the heading in lower case.
Sub-sub-headings appear like sub-headings, except that they are in
italics and not boldface. See examples in this file. No more than 3
levels of headings should be used. Empty lines should be left above and
below each section heading.

\subsection{Text font}

Times or Times New Roman font is used for the main text. Recommended
font size is 11 points. Other font types may be used if needed for
special purposes. If using any non-Unicode fonts, these must be embedded
in the final PDF file.

The \LaTeX{} template can be used either with plain \LaTeX{} or
\XeLaTeX.

\subsection{Figures}

All figures should be centered on the column (or page, if the figure
spans both columns). Figure captions should follow each figure and have
the format given in Fig. \ref{fig:vowels}.

Figures should preferably be line drawings. If they contain grey shades
or colours, it should be checked that they print well on a high-quality
noncolour laser printer.

\begin{figure}[!ht]
\begin{center}
\includegraphics[width=6cm]{./includes/figures/ipa.eps}
\caption{The vowel chart used in the International Phonetic
Alphabet (IPA).}\label{fig:vowels}
\end{center}
\end{figure}

\subsection{Tables}

An example of a table is shown as
Table\textasciitilde{}\ref{tab:decibel}. Somewhat different styles are
allowed according to the type and purpose of the table. Colour should
not be used, but grey shading is allowed. There should be a margin of
6\textasciitilde points (pt) above and below the table.

The caption text may be above or below the table, but this should be
consistent throughout the submission. Left and right indentation of the
caption should be 0.5\textasciitilde cm.

\begin{table}[!ht]
  \begin{center}
  \begin{tabular}{|c|c|}
  \hline
  \rowcolor[gray]{.75}
  ratio    & Decibels\\
  \hline
  $1/1$    & $0$\\
  $2/1$    & $6$\\
  $3.16$   & $10$\\
  $1/10$   & $-20$\\
  $10/1$   & $20$\\
  $100/1$  & $40$\\
  $1000/1$ & $60$\\
  \hline
  \end{tabular}
  \caption{This is an example of a table showing Decibel (dB)
  ratios.}\label{tab:decibel}
  \end{center}
\end{table}

\subsection{Equations}

Equations should be placed on separate lines and numbered. An example of
an equation is \begin{equation}\label{eq:tzero}
T_0 = \frac{1}{f_0}.
\end{equation} Numbers of equations can be on the right or on the left
margin of the text column.

\subsection{Examples}

Examples from other languages can either be presented in the body text,
or, if referred to elsewhere or particularly long and complex, can be
put on a separate, numbered line, as should be done for equations.

\subsection{Phonetic fonts}

We recommend that you use the \textsc{tipa} package for IPA phonetic
symbols. For information about how to access Unicode fonts in the Word
template, see \cite{IPA-SIL} or \cite{IPA-KEYBOARD}. The font you use
must be embedded. Please remember to check this, e.g.~by inspecting the
``Document Properties --- Fonts'\,' in Acrobat Reader.

\subsection{Page numbering}

Page numbers will be added electronically to the document later.
\textit{Please do not add page numbers and please do not insert
any footers or headers!}

\subsection{References}

Please use just the reference number in square brackets. Formulations
with author names like ``\ldots~as Ladefoged \cite{Ladefoged:2003}
showed that \ldots'' are acceptable but not ``as shown in {[}Ladefoged,
3{]}'' or ``as shown in (Ladefoged {[}3{]})''.

References are to be numbered in alphabetical order. Please double-check
the final version of your paper with regard to the correct
correspondence of references to their numbers.

\subsection{Hyperlinks}

Links to URLs or email addresses should be formatted as normal text,
\textit{not} as hyperlinks and not blue or underlined etc. Usually
hyperlinks to web pages are listed in the references section. If
required, line breaks can be placed within URLs after slashes or dashes
(cf.\textasciitilde{}\cite{IPA-SIL,IPA-KEYBOARD}), but doublecheck that
no hyphens are inserted.

\subsection{Footnotes and endnotes}

If footnotes cannot be avoided they should appear as
endnotes.\footnote{This footnote appears here as an endnote.}

\section{Multimedia files}

Multimedia data that are part of the paper are to be embedded in the
submitted PDF; they cannot be submitted as supplementary data. Any
images are to be included in the paper as Figures (see Section 2.3
above). It is the authors' responsibility to check image quality ahead
of submission. Audio examples are to be embedded within the PDF. To do
this, authors can generate the PDF, and then embed the audio files using
Adobe Acrobat Professional, or other software that offers the same
outcome, so that the audio is included in the PDF. The presence of audio
data should be identified in the text.

We encourage authors to illustrate video data using still photographs
from the video, and to include them as figures in the PDF. We cannot
accept embedded video files, but authors are welcome to refer readers to
a URL on the internet where these can be viewed.

\section{PDF details}

PDF files submitted must comply with the following requirements:

\begin{enumerate}
\item all special fonts and symbols must be embedded in the PDF
  file so that correct rendering of the PDF does not depend on the
  fonts installed on the viewer's computer
\item there must be no
  password protection on the PDF file, i.e. PDF files must not be
  protected by PDF security in any way, i.e. content extraction,
  document assembly, high resolution printing etc. must not be
  forbidden
\item PDF files should not contain any colours, hyperlinks,
  multimedia or 3D content, and no JavaScript or forms
\item PDF files
  should be no larger than 5 MB.
\end{enumerate}

\section{Anonymity}

>>>>

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
  a4paper,
  11pt,
  twocolumn]{article}
\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math} % this also loads fontspec
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else
  % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
%------------------------------------------------------------------------------%
% PAPER TEMPLATE FOR ICPHS 2023 Prague                                         %
%                                                                              %
% Original template downloaded from:                                           %
% http://www.icphs2023.org/call-for-papers/                                    %
%                                                                              %
% Reformatted to work with Rmarkdown and R by:                                 %
% Joseph V. Casillas | Rutgers Univesity |11/11/2022                           %
%                                                                              %
% Available for download at:                                                   %
% https://github.com/jvcasillas/icphs2023_rmd_template                         %
%------------------------------------------------------------------------------%

@@

\title{Language Attitudes of Catalan-Spanish Bilinguals Towards Intervocalic Alveolar Fricatives in Catalan}
\author{Please write XXXX instead of the name(s) of the author(s)}
\organization{Please write XXXX instead of the affiliation(s)}
\email{please write XXXX instead of the email address(es)}

@@

\begin{abstract}
Catalan and Spanish are two languages that have coexisted for centuries in Eastern Spain. Thus, there is a bidirectional influence between these languages in all linguistic levels, including in the production of sounds. However, little is known about the perception and the attitudes toward a less Catalan-like production of a given sound. Therefore, the aim of this study is to examine the covert attitudes toward the production of intervocalic /s/, as either [s] or [z] by Catalan-Spanish bilinguals from Catalonia using a matched guise test. Results indicate that phoneme type is not a significant predictor of language attitudes in Catalonia, but other individual social factors, such as mother tongue and province, might be.
\end{abstract}

\keywords{Catalan, Spanish, language attitudes, intervocalic alveolar fricatives.}

@@

Catalan and Spanish are two languages that have coexisted for centuries,
mainly in eastern Spain, along the Mediterranean coast \textbf{(Hualde,
1992)}. It is safe to assume that there is a bidirectional influence
between these languages: in that region, Catalan has influenced Spanish
and Spanish has influenced Catalan in all linguistic levels
\textbf{(Davidson, 2014, 2015)}. A crucial area of cross-linguistic
influence might be the production of sounds, especially in those sounds
that have different distribution in Catalan and Spanish. For example,
previous studies have found that, in Barcelona, the production of
Catalan vowels is different based on the L1 of the speaker, being it
less target-like in L1 Spanish speakers compared to L1 Catalan speakers
\textbf{(CITAR SEBASTIAN GALLES, etc)}.

However, little is known about the perception and the attitudes toward a
less Catalan-like production of a given sound. Therefore, the aim of
this study is to examine the covert attitudes towards the production of
intervocalic /s/, as either {[}s{]} (disfavored by prescriptivist norms)
or {[}z{]} (preferred by prescriptivist norms) by L1 Catalan, L1
Spanish, and Catalan-Spanish bilinguals from Catalonia using a matched
guise test \textbf{(Lambert, 1967)}. The alveolar fricative was chosen
because it exists in both Catalan and Spanish inventories, but their
contextual distribution is different \textbf{(Llisterri, 2023)}.

@@

\textbackslash cite\{hualde\_colina\_2015

While Spanish inventory does not have the phoneme /z/, the sound {[}z{]}
may occur as an allophone of /s/, due to a voice assimilation process.
It is important to mention, though, that it has been claimed that this
voicing is generally not complete, but partial \textbf{(Benet et al.,
2012; Hualde \& Colina, 2015; Hualde et al., 2020)}. There is some
evidence of the influence of Spanish on the production of the Catalan
alveolar fricative, in contexts where, according to prescriptivist rules
\textbf{(Institut d'Estudis Catalans, 2022)}, the {[}z{]} should be
produced, but {[}s{]} is produced instead. This phenomenon has mainly
been attested in those areas where there has been an extended contact
between Catalan and Spanish, such as the so-called ``franja de Aragón''
(borderland between Catalonia and Aragon), some areas of the Valencian
Community, and Barcelona and other urban areas \textbf{(Benet et al.,
2012; Davidson, 2020; Recasens Vives, 1996, p.~268)}. The opposite
pattern has also been observed - that is, the voicing of the alveolar
fricative in the Spanish spoken in Barcelona and Valencia.
Interestingly, although in both Barcelona and Valencia Catalan and
Spanish coexist and influence each other, it seems that Catalan has a
greater influence on Spanish in Barcelona, but Spanish has a greater
influence on Catalan in Valencia \textbf{(Davidson, 2020)}.

@@

The social meaning and the perception of intervocalic {[}z{]} in Spanish
has been studied in a variety of contexts, including both monolingual
\textbf{(Chappell, 2016, 2017, 2021; Chappell et al., 2023; Chappell \&
García, 2017; García, 2015, 2019; Silverstein, 2003)} and bilingual
contexts \textbf{(Davidson, 2014, 2015, 2019; Sinner, 2002)}, despite
having been categorized as a low-salient variant \textbf{(Chappell,
2017, 2021; Chappell et al., 2023; García, 2015)}. With respect to the
former, it has been found that for listeners from Costa Rica and Ecuador
(García, 2019), the covert attitudes associated with intervocalic
{[}z{]} vary significantly by speaker gender. Consequently, male
speakers who use {[}z{]} tend to be positively evaluated, and generally
perceived as nice, local, confident, and masculine. Conversely, female
speakers who use {[}z{]} are only afforded negative associations of
lower education and social class \textbf{(Chappell, 2016; Chappell et
al., 2023; Silverstein, 2003)}. Additionally, this variable has been
found to serve as a marker of regional identity, with voicing being
associated with neighboring Highland Ecuadorian dialects that exhibit
systematic voicing.

With respect to the social meaning and the perception of intervocalic
{[}z{]} in bilingual contexts, this phenomenon has been mainly studied
in Catalan-Spanish contact situations \textbf{(Davidson, 2014, 2015)}.
According to \textbf{Sinner (2002)}, intervocalic {[}z{]} can be
considered a possible linguistic marker of Catalonian Spanish, despite
its relative lack of social awareness \textbf{(Davidson, 2019; Sinner,
2002)}. Additionally, Davidson (2019) found that, in Barcelona, the
indexical field of intervocalic {[}z{]} features exclusively positive
associations with bilingualism and local solidarity, which warrants its
gradual adoption in the speech community, unlike other phonetic markers
of Catalonian Spanish that are the topic of overt social ridicule, such
as lateral velarization. These results are in line with previous
findings on intervocalic {[}z{]} in Catalonian Spanish. For instance,
\textbf{Silverstein (2003)} found that intervocalic {[}z{]} indexes a
local Barcelona identity, and the variant then acquires positive
associations with bilingualism and cultural pride. Overall, these
findings suggest that intervocalic {[}z{]} likely functions as a
sociolinguistic indicator of Catalonian Spanish within the Catalonian
speech community, indexing Catalan-Spanish bilingual membership while
lacking any considerable degree of generalized conscious awareness
\textbf{(Davidson, 2019)}.

@@

Accordingly, and considering that there are no perception studies
focusing on Catalonian Catalan (central variety), the present study aims
to answer the following research questions:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  What covert attitudes are linked to the perception Catalan
  intervocalic /s/ as either {[}s{]} or {[}z{]}?
\item
  What is the relationship between individual social factors and covert
  attitudes towards Catalan intervocalic /s/?
\end{enumerate}

@@

Data were collected using an online survey hosted on Qualtrics that
consisted of two parts. The first part, a demographic questionnaire,
collected information regarding participant age, gender, place of
origin, place of residence, mother tongue, family income, and level of
studies. The second portion of the survey consisted of a perception task
where participants were presented with a series of stimuli containing
the target phoneme contrast (intervocalic voiced and voiceless alveolar
fricative) such as \emph{han tancat les vies estretes} `they closed the
narrow streets', followed by 11 evaluative statements related to social
characteristics such as level of studies, friendliness, social class,
and ethnic origin. These evaluative statements were then followed by a
100-point draggable sliding scale with which participants indicated
their level of agreement or disagreement with the content of the
statement, where 0 was ``totally disagree'' and 100 was ``totally
agree''.

@@

Data were analyzed in R (R Core Team, 2023). For each of the 11
evaluative statements, we created a series of hierarchically nested
linear regression models using the lme4 package (Bates et al., 2015)
with attitudinal sliding scale responses as the dependent variable, and
phoneme, mother tongue, and province as potential predictors. To
determine the best fit model for each statement, we used the anova()
function (Fox \& Weisberg, 2019).

@@

Participant responses to the statements \emph{Parla bé català} `S/he
speaks proper Catalan' and \emph{És d'origen català} `S/he is of Catalan
origin' show similar results, as illustrated in Figure \ref{fig:oc}.
Those who have both Catalan and Spanish as their L1 rate the speaker
more positively regardless of phoneme, compared to those whose L1 is
either Catalan or Spanish. In addition, although it is not statistically
significant, participants give higher ratings to the speaker when they
are listening to those statements with {[}z{]}.

@@

Participant responses to the statements \emph{És simpàtic} `S/he is
nice' and \emph{És de fiar} `S/he is trustworthy' showed similar
results, as illustrated in Figures \ref{fig:fiar}. Those who are from
and live in the province of Barcelona give higher ratings to the
speaker, compared to those who are from and live in Girona.

@@

\section{Discussion}

The aim of this exploration study was to explore the covert attitudes
that are linked to the perception Catalan intervocalic /s/ as either
{[}s{]} or {[}z{]} and to examine the relationship between individual
social factors and covert attitudes towards Catalan intervocalic /s/.
The results suggest that these predictors have a different effect on the
ratings of the speaker depending on the social characteristic under
evaluation.

@@
